---
title: "Aufgabe1"
author: "Max Melchior Lang"
date: "8/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pammtools)
library(tidyverse)
options(pillar.sigfig = 10)
```

# Aufgabe 1
## a)

First a `get_mode` function gets defined, which will get wrapped in the `impute` function.
```{r}
### get_mode function
###
### This function computes the mode that is defined as the  the value that appears most
### often in a set of data values.
### Arguments:
### x: A numeric vector. The vector for which the mode should be
### computed.
### na.rm: A logical of length 1. Should NA's be removed when the mode is
### computed? 
### Returns: A numeric of length one: the computed mode

get_mode <- function(x, na.rm= FALSE) {
  if(na.rm == TRUE){
    x <- na.omit(x)
  }
  
   unique_values <- unique(x)
   unique_values[which.max(tabulate(match(x, unique_values)))]
}

```


```{r}
### impute function
###
### This function imputes missing values in a dataset. For metric variables the 
### median will replace the missing values, while for categorical variables the mode 
### will replace the missing values. This function wraps the previously defined 
### 'get_mode' function. 
### 
### Arguments:
### data: A data frame The dataset for which the missing values should get
### imputed.
###
### Returns: A data frame: The imputed dataset.

impute <- function(data){
  # Input checks
  if(!is.data.frame(data)){
    stop("The Arguments data has to be a data.frame object")
  }
  
  # Iteration over columns and rows
  for (column in colnames(data)){
    # Two cases for imputation
      # Metric variable
      if(is.numeric(data[[column]]) && !(all(na.omit(data[[column]]) %in% 0:1))){
        na_index <- which(is.na(data[[column]]))
        col_median <- median(data[[column]], na.rm = TRUE)
        data[[column]] <- replace(data[[column]], na_index, col_median)
      }
      # Categorical variable
      if(any(is.factor(data[[column]]),is.character(data[[column]]),
           is.logical(data[[column]]), all(na.omit(data[[column]]) %in% 0:1))){
      na_index <- which(is.na(data[[column]]))
      col_mode <- get_mode(data[[column]], na.rm = TRUE)
      data[[column]] <- replace(data[[column]], na_index, col_mode)
    }
  }
  return(data)
}
```

## b)
```{r}
### pairwise_correlation function
###
### This function computes the pairwise correlation (Pearson) for all metric 
### variables in the data set and crosstables for all combinations 
### of categorical variables.
### Pairs of metric and categorical variables are ignored.
### The function returns a list with two elements:
###     1) Correlation Matrix of metric variables in the dataset
###     2) A list of crosstables of all categorical variables in the dataset
###
### Arguments:
### data: A data frame The dataset for which the missing values should get
### imputed.
### Returns: A list of two elements : The list contains 1) a correlation matrix
###                                   all metric variables and 2) a list of
###                                   crosstables of all categorical variables 
###                                   in the dataset.

pairwise_correlation <- function(data){
  # Input checks
  if(!is.data.frame(data)){
    stop("The Argument data has to be a data.frame object")
  }
  
  # Declaration of metric and categorical variables
  metric_cols <- vector()
  catego_cols <- vector()
  
  for(column in 1:ncol(data)){
  # Metric Variables
    if(is.numeric(data[[column]])&& !all(na.omit(data[[column]]) %in% 0:1)){
      metric_cols <- c(metric_cols,colnames(data[column]))
    }
  # Categorical Variables 
    if(any(is.factor(data[[column]]),
         is.character(data[[column]]),
         is.logical(data[[column]]), 
         all(na.omit(data[[column]]) %in% 0:1))){
    catego_cols <- c(catego_cols,names(data[column]))
    }
  }
  
  # Subset of metric and 
  metric_data <- data[metric_cols]
  catego_data <- data[catego_cols]
  
  # Output element metric variables
    # Correlation calculation for metric variables
  cor_matrix <- as.matrix(cor(data[metric_cols], 
                              use= "complete.obs", method ="pearson"))
  
  # Output element categorical variables 
   # Creating Matrix with all combinations for crosstable
  catego_combn <- combn(colnames(catego_data), 2)
   # Creating list with crosstables for each combination,
  data_list <- list()
  for(i in 1:ncol(catego_combn)){
  data_list[[i]] <- prop.table(table(data[[catego_combn[1,i]]],
                                      data[[catego_combn[2,i]]]), margin = 2)
  names(data_list)[i] <- paste(catego_combn[1,i], "vs.", catego_combn[2,i])
  }
  
  # Output (list) of function
  output <- list("correlation_metric_variables"= cor_matrix,
                 "crosstables_catego_variables"= data_list)
  return(output)
}  
```


## c)

```{r}
data("patient", package = "pammtools")
str(patient)

# Imputation

patient <- impute(patient)
# Checking for NA values
na_matrix <- matrix(data = NA, nrow = ncol(patient), ncol= 1)
for (i in 1:ncol(patient)){
  na_matrix[i,1] <- sum(is.na(patient[i]))
}
na_matrix

# Pairwise correlation
patient_no_id <- patient[!(colnames(patient) %in% c("CombinedicuID", "CombinedID"))]
pairwise_correlation(patient_no_id)
```

## d)

The correlation between the age of the patients (`age`) and the time until they 
were released from the hospital (`survhosp`) almost no correlation. The Pearson
correlation coefficient is with `-0.0953` slightly negative.

Taking a look at the crosstable of the diagnose (`DiagID2`) and the categorical variable if or if no the patient survived (`PatientDied`). One can see that (relatively) always more people survived than died for each diagnosis. The highest probability to survive is for an `Orthopedic` and `Metabolic` emergencies, the lowest probability to survive is for a `Sepsis` diagnose.
```{r}
# Age and survhosp
pairwise_correlation(patient_no_id)[[1]][3,2, drop= FALSE]

# DiagID2 and PatientDied 
pairwise_correlation(patient_no_id)[[2]][7]
```


# Session Info
```{r}
sessionInfo()
```


